---
title: "16 Genres"
output: html_notebook
---

The original plot used a data set with 16 genres. Starting afresh I'll derive the list.

```{r setup, include = FALSE}
library(tidyverse)
library(lubridate)

billboard <- readr::read_csv("billboard.csv")
audio_features <- readr::read_csv("audio_features.csv")
```

This is the data used by the plot, just before the section of code labeled "impove gaps" which I gather is used for presentation purposes.
```{r original-plot-data }
original_data <- read_csv("lau_g_by_year2.csv")
```

All the genres used. The last letter for each genre is missing:
```{r list-genres }
knitr::kable(
  original_data %>% select(test.1) %>% distinct()
)
```

# Recreating the list

```{r first-genre}
song_genre <- audio_features %>%
  mutate(
    first_genre = str_extract(spotify_genre, "(?<=')[a-z ]+(?=')")
  ) %>%
  select(song_id, first_genre)
```

We can now add the genre to the billboard data.
```{r join-genre }
billboard_genre <- billboard %>%
  left_join(song_genre, by = "song_id")
```

And follow the same steps used by the author tidyverse style:
```{r tidy-version}
original_tidy <- billboard_genre %>%
  mutate(year = year(mdy(week_id))) %>%
  select(year, first_genre) %>%
  count(year, first_genre) %>% 
  rename(total = "n") %>%
  arrange(desc(total)) %>% 
  group_by(year) %>%
  slice(1:3)%>% 
  filter(!is.na(first_genre) | first_genre == "")
```

I seem to have missed something here. Work through step by step and compare.

```{r import-checkpoint-data }
checkpoint_1_songs  <- read_csv("lau_songs.csv")
checkpoint_2_genres <- read_csv("lau_genres.csv")
checkpoint_3_g_by_year2 <- read_csv("lau_g_by_year2.csv")
```

Recreate the data found in the `songs` variable.
```{r checkpoint-1 }
songs_2 <- left_join(billboard, 
                     audio_features, 
                     by = c('song_id', 'song', 'performer')) %>%
  mutate(date = mdy(week_id)) %>% 
  mutate(year = year(date))

songs_2 <- songs_2 %>% 
  select(year, spotify_genre)

all_equal(songs, songs_2)
```

Recreate the data found in the `genres` variable.
```{r checkpoint-2 }
genres_2 <- transform(songs, test=do.call(rbind, strsplit(spotify_genre, ',', fixed=TRUE)), stringsAsFactors=F)

genres_2 <- genres_2 %>% 
  select(year, test.1)

genres_2$test.1 <- gsub("","",as.character(genres_2$test.1))
genres_2$test.1 <- gsub("'","",as.character(genres_2$test.1))
genres_2$test.1 <- gsub('"',"",as.character(genres_2$test.1))


genres_2$test.1 <- substr(genres_2$test.1, 2, nchar(genres_2$test.1)-1)

all_equal(genres, genres_2)
```

Recreate the data found in the `g_by_year2` variable.
```{r checkpoint-3 }
g_by_year_2 <- genres_2 %>% 
  count(year, test.1) %>% 
  rename(total = "n")

# Top N highest values by group
g_by_year2_2 <- g_by_year_2 %>%                                      
  arrange(desc(total)) %>% 
  group_by(year) %>%
  slice(1:3)


#remove empty cells and NA
g_by_year2_2 <- g_by_year2_2[!(is.na(g_by_year2_2$test.1) | g_by_year2_2$test.1==""), ]

all_equal(g_by_year2, g_by_year2_2)
```

