---
title: "genres variable"
output: html_notebook
date: "2021-10-08"
---

In "songs variable" I worked through the steps to create `songs`, a variable in the original code. The next step is the `genres` variable.

```{r setup, include = FALSE}
library(tidyverse)
library(lubridate)
library(jsonlite)

billboard <- readr::read_csv("billboard.csv")
audio_features <- readr::read_csv("audio_features.csv")
```

I created a "checkpoint" in the original code to save the `songs` and  `genres` variables:

```{r variable-checkpoints }
songs_rds  <- read_rds("songs.rds")
genres_rds <- read_rds("genres.rds")
```

Here is the original code, modified to take the songs_rds and produce the genres_rds:

```{r original-code }
genres.1 <- transform(songs_rds, test=do.call(rbind, strsplit(spotify_genre, ',', fixed=TRUE)), stringsAsFactors=F)

genres.1 <- genres.1 %>% 
  select(year, test.1)

genres.1$test.1 <- gsub("","",as.character(genres.1$test.1))
genres.1$test.1 <- gsub("'","",as.character(genres.1$test.1))
genres.1$test.1 <- gsub('"',"",as.character(genres.1$test.1))


genres.1$test.1 <- substr(genres.1$test.1, 2, nchar(genres.1$test.1)-1)

all_equal(genres.1, genres_rds)
```

We can write our own version of this code if we like. Next up the `g_by_year` and `g_by_year2` variables.

```{r make-lau-genres-function }
lau_genres <- function(songs_var) {
    genres <- transform(songs_var, test=do.call(rbind, strsplit(spotify_genre, ',', fixed=TRUE)), stringsAsFactors=F)
  
  genres <- genres %>% 
    select(year, spotify_genre, test.1)
  
  genres$test.1 <- gsub("","",as.character(genres$test.1))
  genres$test.1 <- gsub("'","",as.character(genres$test.1))
  genres$test.1 <- gsub('"',"",as.character(genres$test.1))
  
  
  genres$test.1 <- substr(genres$test.1, 2, nchar(genres$test.1)-1)
  
  return(genres)
}

lau_genres(songs_rds)
```

This does the same thing, without losing trailing characters.

```{r my-genre-mutate-function }
jr_genres <- function(songs_var) {
  songs_var %>%
    mutate(
      first_genre = str_extract(spotify_genre, "(?<=')[a-z ]+(?=')")
    )
}

jr_genres(songs_rds)
```

A lookup table of the two genre columns would be good too.

```{r lookup-table }
genre_lookup <- tibble(first_genre = jr_genres(songs_rds)$first_genre,
                       test.1 =lau_genres(songs_rds)$test.1)

genre_lookup <- genre_lookup %>% distinct()
```

After a bit of examination (see `genre_lookup.R`) it turns out that there are genres with dashes: 
```{r double-quote-genre }
audio_features %>%
  filter(str_detect(spotify_genre, "-")) %>%
  select(spotify_genre)
```

There are also two genres with a single quote enclosed by double quotes:
```{r}
audio_features %>%
  filter(str_detect(spotify_genre, "\"")) %>%
  select(spotify_genre)
```


Try JSON again?
```{r json-parse}
json_genres <- function(songs_var) {
  songs_var %>%
    mutate(
      json_genre = fromJSON(spotify_genre)
    )
}

json_result <- json_genres(songs_rds)
```



