---
title: "Entry of Genres to Billboard"
output: html_notebook
date: "2021-10-13"
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      messages = FALSE, 
                      include = TRUE,
                      results = "hide")
```

```{r setup, include = FALSE }
library(tidyverse)
library(lubridate)
```

## Data Processing Steps

Import the Billboard & Spotify data
```{r import-data }
billboard <- readr::read_csv("billboard.csv")
spotify <- readr::read_csv("audio_features.csv")
```

Add year column to Billboard 
```{r add-year }
billboard <- billboard %>%
  mutate(
    year = year(mdy(week_id))
  )
```

Retrieve songs by genre
```{r songs_of_genre }
songs_of_genre <- function(genre) {
  genre <- str_c("'", genre, "'")
  
  spotify %>%
    filter(str_detect(spotify_genre, genre))
}
```

Merge billboard & spotify songs
```{r merge_songs }
billboard_with_genre <- function(genre) {
  billboard %>% 
    semi_join(
      songs_of_genre(genre), by = c('song_id', 'song', 'performer')
    )
}
```

Collect the data for the three genres. Note misspelling of "atl hip hop". That's what was used in the original plot.
```{r all-alternative-rock }
alt_rock    <- billboard_with_genre("alternative rock")
alt_metal   <- billboard_with_genre("alternative metal")
alt_hip_hop <- billboard_with_genre("atl hip hop") # Misspelling in data.
```

Find the best years for genres.
```{r top-years }
top_years <- function(data, n) {
  data %>% count(year, sort = TRUE) %>% head(n)
}
```

```{r genre-songs-by-year }
songs_by_year <- function(year_on_charts, bill_data) {
  bill_data %>%
    filter(year == year_on_charts) %>%
    group_by(song, performer, song_id) %>%
    summarise(
      entries = n()
    ) %>%
    arrange(desc(entries)) %>%
    select(-song_id)
}
```

## Alt Rock on the Charts

```{r alt-rock-popularity-plot}
ggplot(alt_rock) +
  geom_freqpoly(aes(year), binwidth = 1) +
  labs(title = "Alt Rock popularity over time",
       x = "Year", y = "Number of Entries") +
  ggrepel::geom_label_repel(aes(x = year, y = n, label = year), 
             data = top_years(alt_rock, 3))
```

### Alt Rock in 1994

```{r alt-rock-popular-songs-1994 }
knitr::kable(
  songs_by_year(1994, alt_rock), 
  col.names = c("Song", "Performer", "Entries")
)
```

### Alt Rock in 1996

```{r alt-rock-popular-songs-1996 }
knitr::kable(
  songs_by_year(1996, alt_rock), 
  col.names = c("Song", "Performer", "Entries")
)
```

### Alt Rock in 1999

```{r alt-rock-popular-songs-1999 }
knitr::kable(
  songs_by_year(1999, alt_rock), 
  col.names = c("Song", "Performer", "Entries")
)
```

## Alt Metal on the Charts

```{r alt-metal-popularity-plot}
ggplot(alt_metal) +
  geom_freqpoly(aes(year), binwidth = 1) +
  labs(title = "Alt Metal popularity over time",
       x = "Year", y = "Number of Entries") +
  ggrepel::geom_label_repel(aes(x = year, y = n, label = year), 
             data = top_years(alt_metal, 3))
```

### Alt Metal in 2002

```{r alt-metal-popular-songs-2002 }
knitr::kable(
  songs_by_year(2002, alt_metal), 
  col.names = c("Song", "Performer", "Entries")
)
```

### Alt Metal in 2003

```{r alt-metal-popular-songs-2003 }
knitr::kable(
  songs_by_year(2003, alt_metal), 
  col.names = c("Song", "Performer", "Entries")
)
```

### Alt Metal in 2004

```{r alt-metal-popular-songs-2004 }
knitr::kable(
  songs_by_year(2004, alt_metal), 
  col.names = c("Song", "Performer", "Entries")
)
```

## Alt Hip-Hop on the Charts

```{r alt-hip-hop-popularity-plot}
ggplot(alt_hip_hop) +
  geom_freqpoly(aes(year), binwidth = 1) +
  labs(title = "Alt hip-hop popularity over time",
       x = "Year", y = "Number of entries") +
  ggrepel::geom_label_repel(aes(x = year, y = n, label = year), 
             data = top_years(alt_hip_hop, 3))
```

### Alt Hip-Hop in 2004

```{r alt-hip-hop-popular-songs-2004 }
knitr::kable(
  songs_by_year(2004, alt_hip_hop), 
  col.names = c("Song", "Performer", "Entries")
)
```

### Alt Hip-Hop in 2006

```{r alt-hip-hop-popular-songs-2006 }
knitr::kable(
  songs_by_year(2006, alt_hip_hop), 
  col.names = c("Song", "Performer", "Entries")
)
```

### Alt Hip-Hop in 2017

```{r alt-hip-hop-popular-songs-2017 }
knitr::kable(
  songs_by_year(2017, alt_hip_hop), 
  col.names = c("Song", "Performer", "Entries")
)
```