---
title: "g_by_year variable"
output: html_notebook
date: "2021-10-09"
---

```{r setup, include = FALSE}
library(tidyverse)
library(lubridate)

billboard <- readr::read_csv("billboard.csv")
audio_features <- readr::read_csv("audio_features.csv")
```

Looking at a short piece of code that comes after the `genres` variable
```{r checkpoints }
genres_rds <- read_rds("genres.rds")
g_by_year_rds <- read_rds("g_by_year.rds")
```

The code from the author's script:
```{r original-code }
g_by_year <- genres_rds %>% 
  count(year, test.1) %>% 
  rename(total = "n")

all_equal(g_by_year, g_by_year_rds)
```

My version, starting from scratch rather than using the checkpoint data:

```{r my-code }
song_and_genre <- audio_features %>%
  mutate(
    first_genre = str_extract(spotify_genre, "(?<=')[a-z ]+(?=')")
  ) %>%
  select(song_id, song, performer, spotify_genre, first_genre)

data.1 <- billboard %>%
  left_join(song_and_genre, by = c('song_id', 'song', 'performer')) %>%
  mutate(year = year(mdy(week_id))) %>%
  select(year, first_genre) %>%
  count(year, first_genre) %>% 
  rename(total = "n") 
```

Using `genre_lookup` we can see that my code includes all the genres in the author's code.

```{r join-with-lookup }
genre_lookup <- make_genre_lookup()
data.1 <- data.1 %>% left_join(genre_lookup, by = "first_genre")
data.1 %>% anti_join(genres_rds, by = "test.1")
```

Sometimes `test.1` has an empty string where `first_genre` is NA. 
```{r empty-1 }
lookup %>% filter(!is.na(test.1), is.na(first_genre))
```

So where test.1 is an empty string, first_genre is `NA`.

```{r na-and-not-empty }
lookup %>% filter(!is.na(test.1), is.na(first_genre), test.1 != "")
```

There were over 1000 rows where where `test.1` has a genre but `first_genre` doesn't. 

But not any more! See `genre_lookup.R` where I've handled all the cases.

Where `spotify_genre` is `[]` test.1 is `""` and first_genre is `NA`

