---
title: "songs variable"
output: html_notebook
date: "2021-10-08"
---

There are three main variables created as the data for the plot:

* `songs`
* `genres`
* `g_by_year2`

In this notebook I hope to understand the first.

```{r setup, include = FALSE}
library(tidyverse)
library(lubridate)

billboard <- readr::read_csv("billboard.csv")
audio_features <- readr::read_csv("audio_features.csv")
```

The original code:
```{r old.songs.variable }
songs.1 <- left_join(billboard, 
                     audio_features, 
                     by = c('song_id', 'song', 'performer')) %>%
  mutate(date = mdy(week_id)) %>% 
  mutate(year = year(date))

songs.1 <- songs.1 %>% 
  select(year, spotify_genre)
```

The above code produces the same output as the rds file so we can use the latter as the basis of the next investigation - the `genres` variable.
```{r compare-to-rds}
songs_rds <- read_rds("songs.rds")
all_equal(songs.1, songs_rds)
```


My new code for producing the data above, plus the first genre extracted from the `spotify_genre` column.
```{r new.songs.variable}
song_genre <- audio_features %>%
  mutate(
    first_genre = str_extract(spotify_genre, "(?<=')[a-z ]+(?=')")
  ) %>%
  select(song_id, spotify_genre, first_genre)

billboard_genre <- billboard %>%
  left_join(song_genre, by = "song_id")

songs.2 <- billboard_genre %>%
  mutate(year = year(mdy(week_id))) %>%
  select(year, spotify_genre)

all_equal(songs.2, songs_rds)
```

Which brings us to the problem to solve - there are rows in my code that aren't in the original. Why is that? Is this why the `count()` in the next step gives us different results?

An `anti_join()` shows the rows produced by my code that aren't in the original. 
```{r find-missing-rows}
songs.2 %>% anti_join(songs.1, by = "spotify_genre")
```

Now we need to find the original rows, since we've selected away a lot of information.
```{r missing-with-all-columns }
songs.3 <- billboard_genre %>%
  mutate(year = year(mdy(week_id)))

songs.3 %>% anti_join(songs.1, by = "spotify_genre")
```

The original join had two more fields, let's add those to the new code:
```{r do-same-join }
song_genre <- audio_features %>%
  mutate(
    first_genre = str_extract(spotify_genre, "(?<=')[a-z ]+(?=')")
  ) %>%
  select(song_id, song, performer, spotify_genre, first_genre)

billboard_genre <- billboard %>%
  left_join(song_genre, by = c('song_id', 'song', 'performer'))

songs.2 <- billboard_genre %>%
  mutate(year = year(mdy(week_id))) %>%
  select(year, spotify_genre)

all_equal(songs.2, songs_rds)
```

OK, that's the answer! I tried to make my join a bit more succinct but lost a few rows along the way.